This is my Assignment for Reproducible Research written as a R Markdown file
=========================

```{r setoptions, echo=TRUE}
```
1. Code for reading in the dataset and/or processing the data


Set working directory, load some data (Motion data - data downloaded on the 21/03/2017)
Step 1: Download the file (copy from dowload directory to Coursera), Set directory.
Step 2: Unzip the File and make new folder called Movement
Step 3: Read and look at the data format and check there are 3 variables (steps, date and interval) with 17,568 observations.
Step 4: Load the knitr and dplyr packages

```{r}
setwd("C:/Users/default.TOSHIBA/Desktop/Coursera")
unzip(zipfile="C:/Users/default.TOSHIBA/Desktop/Coursera/repdata_data_activity.zip",exdir="C:/Users/default.TOSHIBA/Desktop/Coursera/Movement")
data <- read.csv("C:/Users/default.TOSHIBA/Desktop/Coursera/Movement/activity.csv",header=T,sep=",", na.strings ="NA") 
summary(data)
head(data)
tail(data)
dim(data)
library(knitr)
library(dplyr)
opts_chunk$set(echo = TRUE)
```
This shows there are 17,568 observationswith NA values in the steps column

We can ignore the missing values in the dataset, as mentioned in assignment blurb. Could also use complete.cases function

```{r}
data2 <- na.omit(data)
summary(data2)
head(data2)
tail(data2)
dim(data2)
```
Use the function aggregate or group_by to sum the steps per day

```{r}
stepsperday <- aggregate(steps ~ date, data2, sum)
head(stepsperday,20)
tail(stepsperday,20)
dim(stepsperday)
```
2. Now plot a Histogram of the number of steps taken each day

```{r}
hist(stepsperday$steps, col="red", xlab="Number of Steps per Day", main="Histogram of the Number of Steps per Day")
```
3. Calculate the mean and median number of steps taken each day

```{r}
mean(stepsperday$steps)
median(stepsperday$steps)
```
This shows the mean is 10766 and the median is 10765 steps

Now calculate the mean number of steps per interval (similar to above)

```{r}
stepsperinterval <- aggregate(steps ~ interval, data2, mean)
head(stepsperinterval,20)
tail(stepsperinterval,20)
dim(stepsperinterval)
stepsperinterval
```
4. Plot a time series of the average number of steps taken per interval for all days

```{r}
plot(stepsperinterval$interval, stepsperinterval$steps, type='l', col="red", 
     main="Average number of steps for each interval over 53 days", xlab="Interval", 
     ylab="Average number of steps")
```

5. Find the 5-minute interval that, on average, contains the maximum number of steps

```{r}
maxstepsinterval <- which.max(stepsperinterval$steps)
stepsperinterval[maxstepsinterval,]
```
So the interval 835 has the maxiumum number of 206 steps

6. Code to describe and show a strategy for imputing missing data
The number of missing values (NAs)

```{r}
sum(is.na(data))
```
The total number of rows with Nas is 2304

Fill in missing data values----use the mean for that 5-minute interval
Create new dataset 
Loop through the rows of data, find the NA in steps and corresponding interval, 
Identify the mean number of steps for that interval and replace the NA value with the mean value
```{r}
for (i in 1:nrow(data)) {
    if(is.na(data$steps[i])) {
        newdata <- stepsperinterval$steps[which(stepsperinterval$interval == data$interval[i])]
        data$steps[i] <- newdata 
    }
}
```

7. Now sum up the total number of steps per day and plot histogram after missing values are imputed
```{r}
stepsperdayimpute <- aggregate(steps ~ date, data, sum)
head(stepsperdayimpute)
hist(stepsperdayimpute$steps, xlab = "Number of Steps per Day", main = "Histogram of the Number of Steps per Day (Imputed)")
```
Calculate the mean and median number of steps

```{r}
mean(stepsperdayimpute$steps)
median(stepsperdayimpute$steps)
```
The mean is 10766 and the median is 10766. The values have not changed significantly as expected

8. Make a Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
Use the weekday() function and the dataset with the filled-in missing values first.

```{r}
data['typeofday'] <- weekdays(as.Date(data$date))
data$typeofday[data$typeofday  %in% c('Saturday','Sunday') ] <- "weekend"
data$typeofday[data$typeofday != "weekend"] <- "weekday"
head(data,20)
tail(data,60)

data$typeofday <- as.factor(data$typeofday)
dataImputestepsbyinterval <- aggregate(steps ~ interval + typeofday, data, mean)
```
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(ggplot2)
qplot(interval, steps, data=dataImputestepsbyinterval, geom=c("line"), xlab="Interval", 
      ylab="Number of steps", main="") + facet_wrap(~ typeofday, ncol=1)
```

9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report



